StreamFind for mass spectrometry

### 2024-04-26 dev/dev_ideas/ms_overview.Rmd
Major Structuring of Spectra Interface (not complete yet)

Within the MassSpecData class object, standardized methods for plotting data are also available. All plot methods use data that can be obtained with get methods, meaning that the user is able to re-plot the data with a different approach/style. For instance, the `get_tic()` method can be used to obtain the total ion chromatogram (TIC) as a `data.table` with columns *analysis*, *rt* and *intensity*, as shown below for the first analysis. The same TIC data can be directly plotted using the `plot_tic()` method, as shown below.

```{r get-tic-first-analysis}
head(ms$get_tic(analyses = 1))
```

@@

# getter for polarities
ms$get_polarities(analyses = "00_hrms_s_is_pos_cent-r003")

@@

# getter for MS levels. Note that it returns a list not a vector!
ms$get_spectra_levels()

@@

# getter for TIC
head(ms$get_tic(analyses = 2))

# getter for BPC
head(ms$get_bpc())
```

@@

Besides the `get_tic()` and `get_bpc()` methods (shown above). There are four methods to get low level spectra from analyses: `get_spectra()`, `get_eic()`, `get_ms1()` and `get_ms2()`. All methods use the same arguments as the function `make_ms_targets()` to build MS targets and the *analyses* argument to specify which MassSpecAnalysis objects to query the data. All methods return a `data.table`. The documentation in `?MassSpecData` has detailed information about the use of each method. For each get method, a plot method is available with similar interface. Below each get and plot method is demonstrated for the defined targets.

@@

The extracted ion chromatogram (EIC) can be obtained with `get_eic()` and plotted with `plot_eic()` method. The interface is very similar as the previous methods, with analyses and make targets arguments used to specify each analyses to parse the data and which traces to retrieve.

```{r get-eic-01}
ms$get_eic(analyses = 1, mz = mz1, rt = rt1, ppm = 5, sec = 10)
```

@@

While the EIC method collects and sums the traces in each retention time value (i.e., in each scan), the `get_ms1()` and `get_ms2()` methods average the mass traces in a given retention time interval, returning an averaged spectrum representing several scans. The averaging is based on clustering the mass traces according to a given `mzClust` argument as demonstrated below for both MS1 and MS2 data.

```{r get-ms1}
# MS1 averaged spectrum of targets from 247 to 257 (Carbamazepin-d10) and 239 to 245 (Diuron-d6) m/z and +/- 1 second the retention time of each target
ms$get_ms1(analyses = 3, mzClust = 0.005,
  mz = data.frame("mzmin" = c(247, 239), "mzmax" = c(257, 245)),
  rt = data.frame("rtmin" = targets1$rt - 1, "rtmax" = targets1$rt + 1)
) 
```

@@

```{r get-ms2}
# MS2 data of Diuron-d6 for analysis 1
ms$get_ms2(analyses = 1, mz = targets1[2, ])
```

>>>>

Within the MassSpecData class object, standardized methods for plotting data are also available. All plot methods use data that can be obtained with get methods, meaning that the user is able to re-plot the data with a different approach/style. For instance, the `get_spectra_tic()` method can be used to obtain the total ion chromatogram (TIC) as a `data.table` with columns *analysis*, *rt* and *intensity*, as shown below for the first analysis. The same TIC data can be directly plotted using the `plot_tic()` method, as shown below.

```{r get-tic-first-analysis}
head(ms$get_spectra_tic(analyses = 1))
```

@@

# getter for polarities
ms$get_spectra_polarity(analyses = "00_hrms_s_is_pos_cent-r003")

@@

# getter for MS levels. Note that it returns a list not a vector!
ms$get_spectra_level()

@@

# getter for TIC
head(ms$get_spectra_tic(analyses = 2))

# getter for BPC
head(ms$get_spectra_bpc())
```

@@

Besides the `get_spectra_tic()` and `get_spectra_bpc()` methods (shown above). There are four methods to get low level spectra from analyses: `get_spectra()`, `get_spectra_eic()`, `get_spectra_ms1()` and `get_spectra_ms2()`. All methods use the same arguments as the function `make_ms_targets()` to build MS targets and the *analyses* argument to specify which MassSpecAnalysis objects to query the data. All methods return a `data.table`. The documentation in `?MassSpecData` has detailed information about the use of each method. For each get method, a plot method is available with similar interface. Below each get and plot method is demonstrated for the defined targets.

@@

The extracted ion chromatogram (EIC) can be obtained with `get_spectra_eic()` and plotted with `plot_eic()` method. The interface is very similar as the previous methods, with analyses and make targets arguments used to specify each analyses to parse the data and which traces to retrieve.

```{r get-eic-01}
ms$get_spectra_eic(analyses = 1, mz = mz1, rt = rt1, ppm = 5, sec = 10)
```

@@

While the EIC method collects and sums the traces in each retention time value (i.e., in each scan), the `get_spectra_ms1()` and `get_spectra_ms2()` methods average the mass traces in a given retention time interval, returning an averaged spectrum representing several scans. The averaging is based on clustering the mass traces according to a given `mzClust` argument as demonstrated below for both MS1 and MS2 data.

```{r get-ms1}
# MS1 averaged spectrum of targets from 247 to 257 (Carbamazepin-d10) and 239 to 245 (Diuron-d6) m/z and +/- 1 second the retention time of each target
ms$get_spectra_ms1(analyses = 3, mzClust = 0.005,
  mz = data.frame("mzmin" = c(247, 239), "mzmax" = c(257, 245)),
  rt = data.frame("rtmin" = targets1$rt - 1, "rtmax" = targets1$rt + 1)
) 
```

@@

```{r get-ms2}
# MS2 data of Diuron-d6 for analysis 1
ms$get_spectra_ms2(analyses = 1, mz = targets1[2, ])
```
